---
title: "2024_BANA4080_group08_midterm_project"
Author:
  -name: Jay Powell
  -name: Dawson Siebert
  -name: Nikki Carfora
output: html_document
date: "2024-10-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Introduction
###Business Problem

###Our Approach
Mention datasets and explain method

###Proposed Solution
Summarize how the analysis will help the CEO make better decisions

##Packages and Libraries Used
List of libraries Used and purpose of key libraries

```{r data}
library(completejourney)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(lubridate)

transactions <- get_transactions()
promotions <- get_promotions()
products <- products
demographics <- demographics
coupons <- coupons
coupon_redemptions <- coupon_redemptions
campaign_descriptions <- campaign_descriptions
campaigns <- campaigns

household_campaign <- inner_join(demographics, campaigns, by = "household_id")


household_campaign <- inner_join(campaign_descriptions, household_campaign, "campaign_id")
household_campaign <- inner_join(transactions, household_campaign, "household_id", relationship = "many-to-many")


household_campaign <- household_campaign %>%
  mutate(campaign_start_month = month(start_date, label = FALSE)) %>%
  mutate(campaign_end_month = month(end_date, label = FALSE)) %>%
  mutate(transaction_month = month(transaction_timestamp, label = FALSE))
```

```{r testplots}
plot2 <- household_campaign %>%
  ggplot(aes(x = transaction_month, fill = campaign_type)) +
  geom_bar()+
  labs(
    fill = "Campaign Type",
    x = "Month",
    y = "Transactions Completed",
    title = "Popularity of Campaigns by Month",
    subtitle = "ADD SUBTITLE")


plot(plot2)

plot3 <- household_campaign %>%
  semi_join(top5cats, household_campaign, by ="product_category") %>%
  ggplot(aes(x = campaign_type, fill = product_category))+
  geom_bar()+
  scale_y_continuous(labels = scales::comma)+
  labs(
    x = "Campaign Type",
    y = "Transactions Completed",
    fill = "Product Category",
    title = "Popularity of Campaigns by Product Category",
    subtitle = "Using the 5 most purchased product categories, we can determine which Campaign Types influence the most sales.")


plot3

campaigntype <- household_campaign %>%
  ggplot(aes(x = income, fill = campaign_type)) +
  geom_bar() +
  scale_y_continuous() +
  labs(
    x = "Household Income",
    y = "Transactions Completed",
    fill = "Campaign Type",
    title = "Transactions Completed by Household Income and Campaign Type",
    subtitle = "Determining which campaigns are most effective based on household income."
  ) +
  theme_minimal() +  # Optional: Use a minimal theme for a cleaner look
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  

campaigntype

redemption_heatmap <- household_campaign %>%
  group_by(income, household_size) %>%
  summarize(coupon_redemptions = n()) %>%
  ggplot(aes(x = income, y = household_size, fill = coupon_redemptions)) +
  geom_tile() +
  scale_fill_viridis(option ="plasma" , name = "Coupon Redemptions", labels = scales::comma) +
  labs(
    x="Income",
    y="Household Size",
    title="Coupon Redemptions by Income and Household Size")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  

redemption_heatmap

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
