---
title: "MidTerm Project"
Author:
  -name: Jay Powell
  -name: Dawson Siebert
  -name: Nikki Carfora
output: html_document
date: "2024-10-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Introduction
###Business Problem

###Our Approach
Mention datasets and explain method

###Proposed Solution
Summarize how the analysis will help the CEO make better decisions

##Packages and Libraries Used
List of libraries Used and purpose of key libraries

```{r data}
library(completejourney)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(lubridate)

transactions <- get_transactions()
promotions <- get_promotions()
products <- products
demographics <- demographics
coupons <- coupons
coupon_redemptions <- coupon_redemptions
campaign_descriptions <- campaign_descriptions
campaigns <- campaigns

household_campaign <- inner_join(demographics, campaigns, by = "household_id")


household_campaign <- inner_join(campaign_descriptions, household_campaign, "campaign_id")
household_campaign <- inner_join(transactions, household_campaign, "household_id", relationship = "many-to-many")


household_campaign <- household_campaign %>%
  mutate(campaign_start_month = month(start_date, label = FALSE)) %>%
  mutate(campaign_end_month = month(end_date, label = FALSE)) %>%
  mutate(transaction_month = month(transaction_timestamp, label = FALSE))
```

```{r testplot}
plot2 <- household_campaign %>%
  ggplot(aes(x = transaction_month, fill = campaign_type)) +
           geom_bar()+
           labs(
             fill = "Campaign Type",
             x = "Month",
             y = "Transactions Completed",
             title = "Popularity of Campaigns by Month",
             subtitle = "ADD SUBTITLE")
           

plot(plot2)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
